Quality Criteria for SKOS Vocabularies

### 2012-03-20 criteria.tex
fixed intro of 3

We identified quality criteria in existing literature and manually checking existing vocabularies against these criteria. We published our preliminary results online\footnote{\url{https://github.com/cmader/qSKOS/wiki/Quality-Criteria-for-SKOS-Vocabularies}} and requested feedback from experts via public mailing list and informal face to face discussions. Based on the responses we received, we selected a subset \todo{CM}{due to lack of space, we cannot present all criteria. Do we have a good argument, why we present only a subset here? What were our selection criteria?} and formalized them into computable quality filtering function, which, in combination, can serve as quantitative quality indicators for SKOS vocabularies. Since each function operates on a given SKOS vocabulary, they are based on the following SKOS data model definition.

\begin{definition}[SKOS Vocabulary] Let the fully entailed RDFS interpretation~\cite{RDFSEM2012} (enriched by entailment of \texttt{owl:inverseOf} properties) of a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with

%    Based on the RDFS Interpretation $\mathcal{I}$ we define a vocabulary $V = (IR, C, A, LL, LR, SR, AR)$ (cf.~\cite{Hitzler2009}). V is expected to be fully entailed, based on the deduction rules for RDFS entailment defined in \cite{RDFSEM2012}. Let

@@

%\item \(A = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I}) \cup I_{CEXT}(\texttt{skos:Collection}^\mathcal{I})\) be the set of \textbf{aggregation and grouping resources}.

%	\item LL being the set of \textbf{lexical labels}, which are instances of RDF plain literals.
%	\item \(LR \subseteq C \times LL\) being the set of \textbf{lexical relations} associating conceptual resources with lexical labels, e.g., via \texttt{rdfs:label} or subproperties thereof.

@@

%\item \(AR \subseteq C \times A\) being the set of \textbf{aggregation relations} associating conceptual resources with instances of aggregation and grouping resources, hence instances of, e.g., \texttt{skos:inScheme} or \texttt{skos:member} or subproperties thereof.
\end{itemize}

@@

The quality criteria we identified have in common that they are applied on a given SKOS vocabulary and identify a set of resources, which do not fulfill that criterion. Therefore, we can regard a quality criterion as being an abstract filtering function $f:V \rightarrow \mathfrak{P}(IR)$, mapping a given vocabulary to the powerset of its resources. Each quality criterion we found can be implemented as a realization of such an abstract filtering function. Since the return value is a set of resources, statistical analyses can be computed on the number of resources in that set.

% Structure for each metric description
%- origin (refs) and design rationale
%- formal definition that specifies how to compute the metric

In the following, we describe the origins and design rationale for each metric and provide a semi-formal definition that specifies how to implement them in tools. We categorized the metrics \todo{CM,BH}{Briefly describe how we organize/structure the criteria}

@@

\subsection{Labeling and Documentation Criteria}
The criteria defined in this section primarily focus on supporting interaction with human users that either work with the vocabulary itself or interact with a system using the vocabulary for, e.g., search and retrieval purposes.

>>>>

We identified an initial set of quality criteria for SKOS vocabularies by reviewing existing literature and manually examining existing vocabularies. We published our findings online\footnote{\url{https://github.com/cmader/qSKOS/wiki/Quality-Criteria-for-SKOS-Vocabularies}} and requested feedback from experts via public mailing list and informal face to face discussions. Based on the responses we received, we then selected a subset of these criteria and formalized them into computable \textbf{quality checking functions}. Each of these functions operates on a SKOS vocabulary, which we define, for the purpose of this work, as follows:

\begin{definition}[SKOS Vocabulary] Let the fully entailed RDFS interpretation~\cite{RDFSEM2012} (enriched by entailment of \texttt{owl:inverseOf} properties \todo{CM}{why?}) of a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with

@@

@@

\end{itemize}

@@

The commonality of all quality checking functions is that they take a given SKOS vocabulary as input and return a (nested) set, containing the resources that do \todo{CM}{not - sure?} fulfill the quality criterion. Therefore, we can formalize a quality criterion as being an abstract quality checking function $f:V \rightarrow \mathcal{P}(IR)$, mapping a given vocabulary $v \in V$ to the powerset of its resources. Each quality criterion we found can be implemented as a realization of such an abstract filtering function and statistical analysis can be computed on top of the return values.

% Structure for each criterion description
%- origin (refs) and design rationale
%- formal definition that specifies how to compute the metric

In the following, we explain the origins and design rationale for each quality criterion and describe how we implemented the corresponding quality checking function. For better readability we provide semi-formal definitions only where necessary and refer to the online documentation or the implementation source code for further details. We also categorized the identified quality criteria into \emph{labeling and documentation criteria}, \emph{structural criteria}, \emph{Linked Data specific criteria}, and \emph{other criteria}.

@@

\subsection{Labeling and Documentation Criteria}

The criteria defined in this section primarily focus on supporting interaction with human users that either work with the vocabulary itself or interact with a system using the vocabulary for, e.g., search and retrieval purposes.
